---
title: "chi_square_analysis"
author: "Kurai Nhliziyo"
date: "2025-03-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Typically I do a lot of classification analyses. In this project I will do some simple statistical analyses using the hotel sales data. My research question is what factors levels of hotel guest satisfaction. I will focus on the customer review: Excellent, Very Good, Good and Poor. Many of the columns in this particular dataset are categorical. Therefore I will do the chi-square test and test if there is any relationship between two categorical variables.

```{r}
# load libraries
library(tidyverse) # metapackage of all tidyverse packages
library(readxl)  # package to read in excel files
```

## Data Preprocessing



```{r, echo=FALSE}
# load the data
file_path <- 'data/transactions.xlsx'

# read the Excel file, sheet 1
data <- read_excel(file_path, sheet = 1)

# show top 6 rows
head(data)
```
```{r}
# see details
glimpse(data)
```

```{r}
# removing unnecessary columns
data <- data %>% select(-c('cus_name','phone_no','emel','card_no'))

# changing column names
data <- data %>%
  rename(room_type = `room-type`,
        child = `(child)`,
        sales_person = `Sales Person`)
```

```{r}
# check for nulls
sum(is.na(data))
```
```{r}
# check what unique values are in categorical columns
# identify categorical variables (both character and factor)
categorical_vars <- sapply(data, function(col) is.factor(col) | is.character(col))

# show unique values for each categorical variable
lapply(data[, categorical_vars, drop = FALSE], unique)
```

## Research questions
1. Does the room type influence the level of guest satisfaction?
2. Is there a relationship between membership status and level of satisfaction?

## Chi-square test to compare categorical vs categorical variables

```{r}
room_type_table <- table(data$customer_review, data$room_type)
print(room_type_table)
```

```{r}
chisq.test(room_type_table)
```
The first chi-square test resulted in a p-value smaller than 0.05. There is a significant relationship between the room type and level of satisfaction.

```{r}
membership_table <- table(data$customer_review, data$membership)
print(membership_table)
```
```{r}
chisq.test(membership_table)
```

The second test also yielded a small p-value. There is a significant relationship between membership and satisfaction.